---
title: "Introduction"
author: "Steve Cassidy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `austalk` package provides access to the Austalk speech database via the Alveo API. 

[Austalk](https://austalk.edu.au/) is a large state-of-the-art database of modern spoken Australian English from all around the country. Recorded between June 2011 and June 2016, the final database contains full audio-visual data for 861 adult speakers (with age ranging from 18 to 83) from 15 different locations in all Australian states & territories, representing the regional and social diversity and linguistic variation of Australian English, including Australian Aboriginal English.

[Alveo](https://alveo.edu.au/) is a Virtual Laboratory supporting research in speech and 
language and provides a data repository with a rich API.   There is a separate R library
that provides general access to the API from R.  This library builds on that to provide
specific functions relevant to the Austalk collection.

## Installation

```{r}
library(alveo)
library(austalk)
```

## Alveo API

The Alveo API provides access by scripts such as these to the data stored on 
the Alveo system.  Since all data access on Alveo is protected, we need a way
to verify who is accessing data.  On the website you would just log in to Alveo
to verify your identity.  For scripts written in R you need to use an API Key downloaded
from the site.  See [the help pages](http://alveo.edu.au/documentation/getting-access-to-alveo-and-galaxy/whats-an-api-key/) for details of how to find your API key.   This library assumes that you
have downloaded the `alveo.config` file via the **Download API Key** sub-menu and 
saved that file in your home directory. You can find where R thinks your home directory 
is by running:

```r
Sys.getenv("HOME")
```

## Structure of Austalk

The Austalk collection contains data from a large number of speakers. Each speaker
attended up to three **sessions**. Within each session they recorded a number of
**components**, for example reading word lists or being part of a Map Task game.  Within
each component, one or more recordings were made, each recording is an **item**.   Each
item contains recordings of different audio channels and may have associated 
annotation files.

## Information About Speakers

Two functions are provided to retrieve information about speakers in Austalk.  The first,
`speakers()` returns a data frame containing one row per speaker and columns for various 
metadata properties. 

```{r}
spk <- speakers()
names(spk)
```
These properties can be used to filter the list of speakers to match certain criteria. For
example to find females over 75:

```{r}
females <- spk[spk$age>75 & spk$gender=='female',]
females[,c('speaker', 'gender', 'age')]
```

The second function returns information about how many recorded items there are for a given
component for each speaker.  Some speakers are identified in the system but have missing data
for one or more sessions or have incomplete items for a component.   This function can be useful to 
ensure that the speakers you work with have data for the target component.  

```{r} 
spk.items <- speaker_items(c('digits', 'sentences'))
spk.items[1:10,]
```

This result could be merged with the list of speakers obtained earlier:

```{r}
females <- merge(females, spk.items, by='speaker')
females[,c('speaker', 'age', 'digits', 'sentences')]
```

You can see in this example that only those speakers with some data are included in the merged data frame
(speaker 2_1049 is excluded).


## Information about Components

Components are the tasks in Austalk, each consists of a number of items which are individual recordings.
You can find the complete list of component names with the `components` function:

```{r}
components()
```

(Note that this list includes some component names that are not part of the main Austalk collection
those starging with "Emotion" are relevant to another collection and will be removed at some point.)

The component short name can then be used to access a list of prompts used in each component. These 
are the prompts displayed to participants for the read speech components or on-screen instructions for
the spontaneous ones.  For example:

```{r}
prompts('digits')
```

```{r}
prompts('yes-no-closing')
```

## Retrieving Items

Given a list of speakers we can retrieve a list of items for a given component using the `componentItems()` function:

```{r}
items <- componentItems(females, 'digits')
items[1:5,]
```

This returns a data frame containing the speaker, the item identifier and the URL of the item on the
Alveo system. This information can then be used to download data for each item from Alveo using the 
`downloadItems` function.  This function takes a list of channel names or patterns to match
the files that should be downloaded.  A common example would be to download only the 16 bit/16 kHz
version of the speaker headset microphone - these files have the suffix `ch6-speaker16.wav`. The
following example would download these files:

```r
filelist <- downloadItems(items, 'data', 'speaker16')
```
The returned value is a list of local filenames that were downloaded.

The channel list can include more than one pattern, so to download both the audio file and any
associated TextGrid annotations you would use:

```r
downloadItems(items, 'data', c('speaker16', 'TextGrid'))
```

## Contributions

Get items associated with a contribution...



